---
title: "Working with USGS Data and Functions"
author: "Alex Olsen-Mikitowicz"
date: "10/7/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
library(tidyverse)
library(dataRetrieval)
library(dygraphs) #Time series plotting
library(tidyr) 
library(xts) #Time series objects
library(ggthemes)

knitr::opts_chunk$set(echo = TRUE)
```

# Working with USGS data

## Download USGS data using dataRetrieval


### Non function oriented method

```{r,eval=F}
lincoln_q <- readNWISdv('06752260', parameterCd = '00060',
                        startDate = '1975-10-01',endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
  mutate(site = 'lincoln')

#summary(lincoln_q)

elc_q <- readNWISdv('06752280', parameterCd = '00060',
                        startDate = '1975-10-01',endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
  mutate(site = 'elc')


combo_data <- bind_rows(lincoln_q,elc_q)
summary(combo_data)
#summary(elc_q)
```


### Function oriented method


```{r}

q_downloader <- function(site_no = '06752260',
                         site_id = 'lincoln'){
  df <- readNWISdv(site_no, 
                   parameterCd = '00060',
                   startDate = '1950-10-01',
                   endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
    mutate(site = site_id)
  return(df)
}


lincoln_q <- q_downloader(site_no = '06752260',
                          site_id = 'lincoln')
elc_q <- q_downloader(site_no = '06752280',
                      site_id = 'elc')
#
combo_data <- bind_rows(lincoln_q,elc_q)

summary(combo_data)
```


## Plot the discharge data

### Time series (xaxis = date)

```{r}

ggplot(combo_data,aes(x=Date,y=q_cfs,color=site)) + 
  geom_line() 


```

### Density plot 

```{r,  warning=FALSE}
ggplot(combo_data,aes(x=q_cfs,color=site)) +
  geom_density() +
  scale_x_log10()

```


### Interactive graphs 

```{r,eval=F}
#Structore examination
#str(combo_data)
wide_data <- combo_data %>%
  select(-agency_cd,-site_no,-q_cd) %>%
  spread(key=site,value=q_cfs)

wide_xts <- xts(wide_data %>%
                  select(elc,lincoln),
                order.by = wide_data$Date)

dygraph(wide_xts)
```


### Function for interactive graph

```{r}

xts_maker <- function(df){
  wide_data <- df %>%
    select(-agency_cd,-site_no,-q_cd) %>%
    spread(key=site,value = q_cfs)

  wide_xts <- xts(wide_data %>%
                 select(-Date),
                order.by = wide_data$Date)
  
  return(wide_xts)
  
}


wide_xts <- xts_maker(combo_data)
wide_elc <- xts_maker(elc_q)

dygraph(wide_xts)
dygraph(wide_elc)
```


# In class work

## Write a function that makes ggplot time series graph of discharge

Here I want you to write your own function that adds custom 
themes to your normal ggplot call (so you don't have to write
out theme_few and theme(legend.position = ...)) over and over again. You may need to load the library `ggthemes`. 

This function call should generate a clean time series plot 
with a single command and a single option (which data frame to 
plot). Please plot the `combo_data` dataset here.


```{r}

q_grapher <- function(data_to_plot){
  p1 <- ggplot(data_to_plot,
               aes( x = Date, y = q_cfs, color = site))+
    geom_line()+
    theme_few()

  return(p1)
}


q_grapher(combo_data)
```


## Download discharge data

Here I want you to download data from any 2 other USGS site in the world using the `q_downloader` function we already wrote. If this 
function works, great! If not, you will need to investigate why? 
Even if your function call works, why is my function "brittle?"

Hint: hardcoding `q_cfs = X_00060_00003` could be dangerous

```{r}

q_downloader <- function(site_no,
                         site_id){
  df <- readNWISdv(site_no, 
                   parameterCd = '00060',
                   startDate = '1950-10-01',
                   endDate = '2019-10-01') %>%
  rename(q_cfs = 4,
         q_cd = 5) %>%
    mutate(site = site_id)
  return(df)
}

q_riog_cerro <- q_downloader(site_no = "08263500",
                             site_id = "cerro_upper")

q_riog_embudo <- q_downloader(site_no = "08279500",
                             site_id = "embudo_lower")

combo_data_rio <- bind_rows(q_riog_cerro , q_riog_embudo)

```


## Joining your two datasets

Above, I combined datasets by stacking them on top of each other 
(bind_rows), but then we had to `spread` the data anyways because
`xts` and `dygraphs` prefer wide datasets. Here I want you 
to write a function that joins your two USGS datasets by calling one of the family of `dplyr::join`. 

Hint: be careful of having two `site` columns. 

```{r}
#Here is where matt and I got stuck
#Primer question 1: can you explain what an expression is and how an expression that starts with ~ is different
#Previously when piping I would call a function on a lower line and have to use (.) as a placeholder for the dataframe or object, from the primer I can run code that looks like this: parameters %>% pmap(rnorm)
#Primer Q 2: what is the difference between class and attribute
#Primer Q 3: Can we talk about the difference between lists, dataframes, and tibbles
#Primer Q 4: Difference between recursive vector and non-recursive vector - difference between lists and vectors
#Primer Q5: basically what is the nature of column lists


xts_maker <- function(df){
  wide_data_rio <- df %>%
    select(-agency_cd,-site_no,-q_cd) %>%
    spread(key=site,value = q_cfs)

  wide_xts_rio <- xts(wide_data_rio %>%
                 select(-Date),
                order.by = wide_data_rio$Date)
  
  return(wide_xts_rio)
  
}


wide_xts_rio <- xts_maker(combo_data_rio)


```


## Plotting the data

My function above `xts_maker` merely preps the data for plotting 
a dygraph. Here, I want you to make a similar function that preps *and* plots a dygraph of your joined data. 


```{r}

xts_plotter <- function(df){
  wide_data_rio <- df %>%
    select(-agency_cd,-site_no,-q_cd) %>%
    spread(key=site,value = q_cfs)

  wide_xts_rio <- xts(wide_data_rio %>%
                 select(-Date),
                order.by = wide_data_rio$Date)
  
  dygraph(wide_xts_rio)
  
}


xts_plotter(combo_data_rio)


```


